\chapter{Docker}
\label{ch:docker}
\newthought{Docker is} ``an open platform for developing, shipping, and running applications''. It is a convenient way for us to provide all of the dependencies and the latest release source code so that we can use the ginan tool kit straight out of the box.
%
In order for this to work, we will first need to install the docker engine onto our local machine. If we are running a different operating system instructions on how to install docker can be found at \href{https://docs.docker.com/get-docker/}{docker desktop downlod link}, these also include alternative methods of installing on ubuntu and has links to recommended best practices.
%
To find more information on docker have a look at the \href{https://docs.docker.com/get-started/}{getting started guide} provided by docker.
%
\section{Ubuntu Docker dependency installation guide}
If we are running ubuntu, we can install a docker engine. A summary of the commands to download and install docker involve setting up the ubuntu repository system to link with the docker repsotory are given below.

\begin{lstlisting}[language=bash]
$ sudo apt -y update
$ sudo apt -y install \
    apt-transport-https \
    ca-certificates \
    curl \
    gnupg \
    lsb-release
\end{lstlisting}

Add the dockers official GPG key:
\begin{lstlisting}[language=bash]
$ curl -fsSL https://download.docker.com/linux/ubuntu/gpg | sudo gpg --dearmor -o /usr/share/keyrings/docker-archive-keyring.gpg
$ echo \
"deb [arch=amd64 signed-by=/usr/share/keyrings/docker-archive-keyring.gpg] https://download.docker.com/linux/ubuntu $(lsb_release -cs) stable" | sudo tee /etc/apt/sources.list.d/docker.list > /dev/null
\end{lstlisting}
%
Then update the repository management system and install the packages.
%
\begin{lstlisting}[language=bash]
$ sudo apt-get update
$ sudo apt-get install docker-ce docker-ce-cli containerd.io
\end{lstlisting}

Verify that the Docker engine is installed correctly by running the \emph{hello-world} image.
\begin{lstlisting}[language=bash]
$ docker run hello-world
\end{lstlisting}

Then we will need to change the ownership:
\begin{lstlisting}[language=bash]
$ sudo usermod -aG docker root
$ sudo usermod -aG docker ${USER}
\end{lstlisting}
You will need to log out and back in for the permissions to take effect.

\section{Using Docker}

Once we have docker installed on our local machine we will need to download our image for ginarn:
\begin{lstlisting}[language=bash]
$ docker pull gamichaelmoore/ginarn-base-ubuntu-20.04
\end{lstlisting}
Then we can run the image as follows:
\begin{lstlisting}[language=bash]
$ docker run -it -v /data:/data gamichaelmoore/ginarn-base-ubuntu-20.04 bash
\end{lstlisting}
This gives us a run-time environment where the dependencies of ginan are installed.
Here, the \texttt{-v} option mounts a volume inside the docker instance at \texttt{/data},
which maps to the \texttt{/data} folder of the host machine. This way this folder
can be shared between the host and the container, and the results can persist.

You should now see a \texttt{bash} prompt running inside the docker container.

\section{Building ginan}

\begin{subsection}{Note}
The instructions here are for separate \texttt{pea} and \texttt{pod} git repositories.
The docker image for the consolidated repository is still a work-in-progress
\end{subsection}

To clone the latest version from the git repositories, run the following from inside the container:
\begin{lstlisting}[language=bash]
mkdir -p /data/acs
pushd /data/acs
# clone pea and pod
# use your own credentials to clone pea
git clone https://git@bitbucket.org/geoscienceaustralia/pea.git
git clone https://anonymous:pipelines@bitbucket.org/geoscienceaustralia/pod.git
popd
\end{lstlisting}

Now, with all the dependencies set up already, we can build the executables:

\begin{lstlisting}[language=bash]
# building pea and pod
cd /data/acs/pea/cpp
mkdir build
cd build
cmake ..
cmake --build /data/acs/pea/cpp/build --target pea
cd /data/acs/
cd pod
mkdir build && cd build
cmake ..
make
export PATH=$PATH:/data/acs/pod/bin
\end{lstlisting}

This docker image comes with a pre-built conda environment that enables plotting. To use it:
\begin{lstlisting}[language=bash]
conda activate gn37
\end{lstlisting}

\section{Keeping a container running}
If we instantiate a container this way, our session will finish when we quit the bash prompt.
The changes we make to the container will also be lost, except the changes that persist outside
of the container, that is, the \texttt{/data} folder in our example.

Therefore, it is sometimes useful to keep a container running, and connect to it and disconnect
from it as needed.

To start up a docker container in the disconnected mode, run:
\begin{lstlisting}[language=bash]
docker run -d -v /data:/data gamichaelmoore/ginarn-base-ubuntu-20.04 sleep infinity
\end{lstlisting}
we can verify that the container is running in the background:
\begin{lstlisting}[language=bash]
docker ps
\end{lstlisting}
This will show a container ID. Docker conveniently also provides an alias as a ``name``.
We can start a new bash shell inside the container by:
\begin{lstlisting}[language=bash]
docker exec -it <name> bash
\end{lstlisting}
where \texttt{<name>} is the name or ID of the running docker container.



